
---
Name: Ryan Chui
Class: Stat 153
title: "HW6"
output: html_document
---

2 (a) Consider the data xt = sin(πt2/256) for t = 0, 1, . . . , 127. Plot the data. Also plot the magnitude (absolute) of the DFT coefficients b1, . . . , b64. Comment on the two plots.
```{r}
xs <- seq(0,127)
wave.1 <- sin((pi*xs^2)/256)
par(mfrow = c(1, 2))
plot(xs,wave.1,type="l",ylim=c(-1,1), xlab="time",ylab="f(t)")
abline(h=0,lty=3)

wave.1.abs <- abs(sin((pi*xs^2)/256))
par(mfrow = c(1, 2))
plot(xs,wave.1.abs,type="l",ylim=c(-1,1),xlab="time",ylab="f(t)")
abline(h=0,lty=3)
```

```{r}
xs <- seq(0,127)
wave.2 <- sin((pi*xs^2)/512)
par(mfrow = c(1, 2))
plot(xs,wave.1,type="l",ylim=c(-1,1), xlab="time",ylab="f(t)")
abline(h=0,lty=3)

wave.3 <- sin((pi*xs^2)/1024)
par(mfrow = c(1, 2))
plot(xs,wave.2,type="l",ylim=c(-1,1),xlab="time",ylab="f(t)")
abline(h=0,lty=3)
```

4. Consider the following seasonal AR model:
(1 − φB)(1 − ΦBs)Xt = Zt,
where {Zt} is white noise and |φ| < 1, |Φ| < 1.
(a) Calculate the spectral density of {Xt}.
```{r}


```
(b) Plot the spectral density for φ = 0.5, Φ = 0.9, σz^2 = 1 and s = 12
```{r}
# Set seed for reproducibility
set.seed(5532)
# Generate a SARIMA(1,0,1)(1,0,0)[12] 
# spec.ar estimate spectral density of a time series from AR fit
a = spec.ar(arima.sim(list(order = c(1,0,0), ar = 0.5, seasonal = c(1,0,0), ar = 0.9, period=12), sd = 1, n = 100))
# Plotting Spectral Densities
# plot.spec.phase(a)

```

(c) Also plot the spectral density for the AR(1) process (1 − 0.5B)Xt = Zt and the seasonal AR(1) process (1 − 0.9B12)Xt = Zt.
```{r}
# Set seed for reproducibility
set.seed(5532)

psi <- function(z) 1/(1 - 0.5*z)
f <- function(x) abs(psi(exp(2i*pi*x)))^2
plotPGram <- function(data, smooth) {
  k <- kernel("daniell", if (smooth) floor(sqrt(length(data))) else 0)
  title <- sprintf("Raw periodogram for %d samples", length(data))
  if (smooth)
  title <- sprintf("Smoothed periodogram for %d samples", length(data))
  p <- spec.pgram(data, k, taper=0, log="no", ylim=c(0,40), main=title)
  grid <- (0:50) / 100
  lines(grid, f(grid), lty=2)
  df <- p$df
  U <- df / qchisq(0.025, df)
  L <- df / qchisq(0.975, df)
  len <- length(p$spec)
  idx <- round(len/5) # Spectral density at 0.1
  # Return a confidence interval
  c(p$spec[idx] * L, p$spec[idx] * U)
}

q4 <- function(n) {
  x <- arima.sim(model=list(ar=0.5), n)
plotPGram(x, F)
}

q4(30)

# Set seed for reproducibility
set.seed(5532)
q4seasonal <- function(n) {
  x <- arima.sim(model = list(seasonal = c(1,0,0), ar = 0.9, period=12), n)
plotPGram(x, F)
}
q4seasonal(30)
```

(d) Compare and comment on the different plots.




Q6
Consider the stationary Autoregressive process:
Xt − 0.99Xt−3 = Zt
where {Zt} is white noise.
```{r}
# Compute Explicitly
omega = seq(from = 0,to = 1/2, length = 110)
a = -2 * pi * 1i * omega
fx = 1 / Mod(1 - 0.99*exp(3*a))^2
plot(omega, fx , type = "l",log = "y")
omega[which.max(fx)]
abline(v = omega[which.max(fx)], col = "blue")
```

b) Does the spectral density suggest that the sample paths of {Xt} will exhibit approximately oscillatory behaviour? If yes, then with what period?
```{r}
# z = exp(-2 * pi * 1i * w)
p = (1 / 0.99)^(1/3)
w = log(p) / (-2 * pi * 1i)
w

# Scale the closest points on unit circle
z = p / Mod(p)
w = log(z) / (-2 * pi * 1i)
w

```

Yes, the spectral density suggest that the sample paths of {Xt} will exhibit approximately oscillatory behaviour. fx(w) is large for w = 0

(c) Simulate a sample of size 100 from this model. Plot the simulated data. Does this plot support the conclusion of part (b)?
```{r}
# Recall AR(3) model
par(mfrow=c(2,1))
set.seed(12345) # for reproducibility
plot(ts(arima.sim(list(ar = c(0,0,0.99)), n = 100)))
```

(d) Compute the spectral density of the filtered process:
Yt = (Xt−1 + Xt + Xt+1)/3                         (1)
How does the spectral density of {Yt} compare to that of {Xt}?
```{r}
# Compute Explicitly
w = seq(from = 0, to = 1/2, length = 110)
# Let sigma2 = 1
fx = 1 / 9 * (3 + 4 * cos(2 * pi * w) + 2*cos(4 * pi * w))
plot(omega, fx , type = "l", log = "y")
omega[which.max(fx)]
abline(v = omega[which.max(fx)], col = "blue")
```

(e) From the simulated sample from {Xt} in part (c), perform the averaging as in (1) to obtain a simulated sample from {Yt}. Plot this sample. Does this plot support the spectral density plot in part (d)?
```{r}
par(mfrow=c(2,1))
plot(ts(arima.sim(list(ar = c(0,0,0.99)), n = 100)))
```

7. Without using the arima.sim() function in R, simulate n = 400 observations from the multiplicative seasonal ARMA model given by the difference equation: (1 − 0.5B)(1 − 0.7B 12)Xt = Zt where {Zt} is white noise. Plot the sample autocorrelation function of the simulated observations and compare it with the true acf of the process.
```{r}
library('astsa')
set.seed(12345)
# fit a seasonal ARIMA(1, 0, 0) x (1, 0, 0)12 model to the series in x
v = rnorm(400,1,1)
x = cumsum(v)  # x is a random walk with drift = 1 
sarima(x,1,0,0,1,0,0,12,  details=FALSE)
# Now, forecast
sarima.for(x,400,1,0,0,1,0,0,12)
acf2(x, 30)  # computes to lag sqrt(n)+10 ; 
             # specify the number of lags... 20

# A seasonal AR(1) model has significant lags at 1, 12 and 13, with an increasing trend before lag 12, sharply cut off at lag 13.
ar_pacf<-ARMAacf(ar = c(.5,0,0,0,0,0,0,0,0,0,0,.7,-.30),lag.max=30)
plot(ar_pacf,type='h')

#sim.ar<-arima.sim(list(ar=c(0.4,0.4)),n=1000)
# acf(sim.ar,main="ACF of AR(2) process")
```

8. Consider the first dataset, q1train.csv, which is uploaded to piazza. Remove the trend and seasonality by differencing first with order 52 and then a usual differencing. Call the resulting dataset xt, t = 1, . . . , n to which a stationary model can be fit.
(a) Estimate the spectral density of {Xt} nonparametrically from the data {xt}
```{r}
library(TSA)
library(sapa)
q1train <- read.csv("~/Downloads/q1train.csv")
train_diff = diff(diff(q1train[,2],52))
## use SDF to estimate the process (cross) spectral density function via nonparametric models
SDF(train_diff)
```

(b) Fit a reasonable stationary model to {xt} and estimate the spectral density of {Xt} by the spectral density of the fitted model
Consider the following model
∇Yt=∑j=1m[Ajcos(2πfjt)+Bjsin(2πfjt)]+et
where m is the number of key frequencies, Aj and Bj are unknown constants, fj’s are the key frequencies, {et} is normal white noise.
```{r}
spec_diff = periodogram(train_diff)
# estimated spectral density greater than 9
key_freq <- spec_diff$freq[which(spec_diff$spec > 3^2)]
t <- 1:length(train_diff)
harmonics <- do.call(cbind, lapply(key_freq, function(freq){
  cbind(cos(2 * pi * freq * t), sin(2 * pi * freq * t))
}))
reslm <- lm(train_diff ~ harmonics)
```

(c) Plot the two estimates of the spectral density on the same plot. Commment on the two
plots.
```{r}
plot(t, train_diff, type="l")
lines(fitted(reslm)~t, col=4, lty=2)
```
